### 2022-03-07 _900_output/paper/paper.tex
Add skeleton for paper

\RequirePackage[thmmarks]{ntheorem}
\makeatletter
\renewtheoremstyle{plain} 
  {\item[\hskip\labelsep \theorem@headerfont ##1\ \textup{##2}\theorem@separator]} 
  {\item[\hskip\labelsep \theorem@headerfont ##1\ \textup{##2}\ (##3)\theorem@separator]}
\makeatother

%%% save \document and \arabic to be reinstated after loading the class
\let\latexdocument\document
\let\latexarabic\arabic

\documentclass[manuscript]{biometrika}
%\documentclass[supplementary,lineno]{biometrika}

%%% reinstate the original \document and \arabic
\let\document\latexdocument
\let\arabic\latexarabic

%%% make \rm into a no-op
\def\rm{}

\usepackage{amsmath}
%\usepackage{graphics}

%% Please use the following statements for
%% managing the text and math fonts for your papers:
\usepackage{times}
%\usepackage[cmbold]{mathtime}
\usepackage{bm}
\usepackage{natbib}
\usepackage{dsfont}

\graphicspath{{./art/}}

\usepackage[plain,noend]{algorithm2e}

\makeatletter
\renewcommand{\algocf@captiontext}[2]{#1\algocf@typo. \AlCapFnt{}#2} % text of caption
\renewcommand{\AlTitleFnt}[1]{#1\unskip}% default definition
\def\@algocf@capt@plain{top}
\renewcommand{\algocf@makecaption}[2]{%
  \addtolength{\hsize}{\algomargin}%
  \sbox\@tempboxa{\algocf@captiontext{#1}{#2}}%
  \ifdim\wd\@tempboxa >\hsize%     % if caption is longer than a line
    \hskip .5\algomargin%
    \parbox[t]{\hsize}{\algocf@captiontext{#1}{#2}}% then caption is not centered
  \else%
    \global\@minipagefalse%
    \hbox to\hsize{\box\@tempboxa}% else caption is centered
  \fi%
  \addtolength{\hsize}{-\algomargin}%
}
\makeatother

%%% User-defined macros should be placed here, but keep them to a minimum.
\def\Bka{{\it Biometrika}}
\def\AIC{\textsc{aic}}
\def\T{{ \mathrm{\scriptscriptstyle T} }}
\def\v{{\varepsilon}}

\addtolength\topmargin{35pt}
\DeclareMathOperator{\Thetabb}{\mathcal{C}}

\begin{document}

% \jname{Biometrika}
% %% The year, volume, and number are determined on publication
% \jyear{2021}
% \jvol{103}
% \jnum{1}
% %% The \doi{...} and \accessdate commands are used by the production team
% %\doi{10.1093/biomet/asm023}
% \accessdate{Advance Access publication on 31 July 2018}

% These dates are usually set by the production team
% \received{2 January 2017}
% \revised{8 June 2021}

%% The left and right page headers are defined here:
% \markboth{P. Fearnhead et~al.}{Biometrika style}

%% Here are the title, author names and addresses
\title{Estimating the size of a population through repeated sampling: a new view on capture-recapture procedures}

\author{G. KAUERMANN, N. SAPARGALI}
\affil{Department of Statistics, Ludwig-Maximilians-Universität München, 80539 Munich, Germany
\email{nurzhan.sapargali@stat.uni-muenchen.de} \email{goeran.kauermann@stat.uni-muenchen.de}}

\maketitle

\begin{abstract}
There should be a single paragraph summary which should not contain formulae or symbols, followed by some key words in alphabetical order.  Typically there are 3--8 key words, which should contain nouns and be singular rather than plural.  The summary contains bibliographic references only if they are essential.  It should indicate results rather than describe the contents of the paper: for example, `A simulation study is performed' should be replaced by a more informative phrase such as `In a simulation our estimator had smaller mean square error than its main competitors.'
\end{abstract}

\begin{keywords}
Capture-recapture estimator; Inclusion probability; Population size.
\end{keywords}

\section{Introduction}

\section{Model}
Consider a closed population of $N$ individuals $(i = 1,\ldots, N)$ from which we draw $T$ samples $(t = 1,\ldots, T)$ using some sampling scheme. Define $S_t$ to be the set of individuals included into sample $t$ and $n_t$ to be the size of $S_t$. Furthermore, let $\pi_i$ and $X_{i(T)}$ denote respectively the inclusion probability of $i$ and the count of $i$ being included into a sample after $T$ draws
\begin{equation*}
\pi_i = \text{pr}(i \in S_t), \quad X_{i(T)} = \sum_{t=1}^T \mathds{1}\{i \in S_t\};
\end{equation*}
If $n_t = n$ for all $t$ and $\pi_i = n / N$ for all $i$, then $X_{i(T)}$ iid

\end{document}

### 2022-03-08 _900_output/paper/paper.tex
Some further changes to paper body

@@
 \affil{Department of Statistics, Ludwig-Maximilians-Universität München, 80539 Munich, Germany
-\email{nurzhan.sapargali@stat.uni-muenchen.de} \email{goeran.kauermann@stat.uni-muenchen.de}}
+\email{goeran.kauermann@stat.uni-muenchen.de} \email{nurzhan.sapargali@stat.uni-muenchen.de}}
 
@@
 
-\section{Model}
-Consider a closed population of $N$ individuals $(i = 1,\ldots, N)$ from which we draw $T$ samples $(t = 1,\ldots, T)$ using some sampling scheme.
-Define $S_t$ to be the set of individuals included into sample $t$ and $n_t$ to be the size of $S_t$.
-Furthermore, let $\pi_i$ and $X_{i(T)}$ denote respectively the inclusion probability of $i$ and the count of $i$ being included into a sample after $T$ draws
+\section{Model and likelihood}
+Our starting point is a basic capture-recapture model with equal capture probabilities.
+Consider a closed population $\mathcal{P} = \{1,\ldots,N\}$ from which we draw $T > 1$ independent samples of size $n$ using some sampling scheme.
+Define $S_t$ to be the set of individuals included into sample $t$ and $\pi_i = \text{pr}(i \in S_t)$ to be the inclusion probability of $i$.
+Let $X_{i(T)}$ denote the frequency count of $i$ being included into $T$ samples
 \begin{equation*}
-\pi_i = \text{pr}(i \in S_t), \quad X_{i(T)} = \sum_{t=1}^T \mathds{1}\{i \in S_t\};
+X_{i(T)} = \sum_{t=1}^T \mathds{1}\{i \in S_t\};
 \end{equation*}
-If $n_t = n$ for all $t$ and $\pi_i = n / N$ for all $i$, then $X_{i(T)}$ iid
-
+If $\pi_i = n / N$ for all $i$, then $X_{1(T)}, \ldots, X_{N(T)}$ are modeled as identically and independently distributed binomial variables
+\begin{equation*}
+\text{pr}(X_{i(T)} = x_i) = \binom{T}{x_i} \left(\frac{n}{N}\right)^{x_i} \left(1 - \frac{n}{N}\right)^{T - x_i}
+\end{equation*}
+However, since we only observe $X_{i(T)} > 0$, the above distribution must be truncated at zero.
+For convenience purposes, it is useful to decompose $\mathcal{P}$ into $S^{in}_T = \{i \in \mathcal{P}: X_{i(T)} > 0\}$ and $S^{out}_T = \{i \in \mathcal{P}: X_{i(T)} = 0\}$.
+The likelihood function is then
 \end{document}

### 2022-03-08 _900_output/paper/paper.tex
More additions to the paper

@@
 Our starting point is a basic capture-recapture model with equal capture probabilities.
-Consider a closed population $\mathcal{P} = \{1,\ldots,N\}$ from which we draw $T > 1$ independent samples of size $n$ using some sampling scheme.
+Consider a closed population $\mathcal{P} = \{1,\ldots,N\}$ from which we independently draw $T > 1$ samples of size $n$ using some sampling scheme.
 Define $S_t$ to be the set of individuals included into sample $t$ and $\pi_i = \text{pr}(i \in S_t)$ to be the inclusion probability of $i$.
@@
 \begin{equation*}
-\text{pr}(X_{i(T)} = x_i) = \binom{T}{x_i} \left(\frac{n}{N}\right)^{x_i} \left(1 - \frac{n}{N}\right)^{T - x_i}
+\text{pr}(X_{i(T)} = x_i; T, N, n) = \binom{T}{x_i} \left(\frac{n}{N}\right)^{x_i} \left(1 - \frac{n}{N}\right)^{T - x_i}
 \end{equation*}
 However, since we only observe $X_{i(T)} > 0$, the above distribution must be truncated at zero.
-For convenience purposes, it is useful to decompose $\mathcal{P}$ into $S^{in}_T = \{i \in \mathcal{P}: X_{i(T)} > 0\}$ and $S^{out}_T = \{i \in \mathcal{P}: X_{i(T)} = 0\}$.
-The likelihood function is then
+For convenience purposes, it is also useful to partition $\mathcal{P}$ into $D = \cup_{t=1}^T S_t = \{i \in \mathcal{P}: X_{i(T)} > 0\}$ and $U = \mathcal{P} \setminus D = \{i \in \mathcal{P}: X_{i(T)} = 0\}$.
+Accordingly, population size can be expressed as a sum of cardinalities of $D$ and $U$, i.e.
+$N = N_D + N_U$, where the first term is known and the latter term is to be estimated.
+The resulting likelihood of the data is
+\begin{equation*}
+\mathcal{L}(N_U) = \prod_{i \in D} \frac{\text{pr}(X_{i(T)} = x_i; T, N, n)}{1 - \text{pr}(X_{i(T)} = 0; T, N, n)} = \prod_{i \in D} \binom{T}{x_i} \frac{n^{x_i} (N_D + N_U - n)^{T-x_i}}{(N_D + N_U)^{T} - (N_D + N_U - n)^T} \\
+\end{equation*}
+Setting $T = 2$ and recognizing that $\sum_{i \in D} x_i = nT$ yields the following maximum likelihood estimator of $N_U$
+\begin{equation*}
+\widehat{N}_U = \frac{(n - N_D)^2}{2n - N_D}
+\end{equation*}
+Thus, the population size is estimated by
+\begin{equation} \label{eq:1}
+\widehat{N} = N_D + \widehat{N}_U = \frac{n^2}{2n - N_D}
+\end{equation}
+Equation \eqref{eq:1} is the special case of the Lincoln-Peterson estimator with fixed $n$ for each sample draw \citep{Pollock:1990}.
+
+Assume now that $\pi_i$ varies across population units meaning that \eqref{eq:1} is no longer applicable.
+Let us consider the problem from the Bayesian perspective and assume a beta prior for inclusion probabilities with hyperparameters $\alpha$ and $\beta$
+\begin{equation*}
+f_{\pi_i}(\upsilon; \alpha, \beta) = \frac{\upsilon^{\alpha - 1}(1 - \upsilon)^{\beta - 1}}{\mathrm{B}(\alpha, \beta)}, \quad \mathrm{B}(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)};
+\end{equation*}
+where $\Gamma(x)$ is the gamma function. From Bayes' theorem, we have
+\begin{equation*}
+f_{\pi_i \mid X_{i(T)}} (\upsilon \mid x_i) = \frac{\text{pr}(X_{i(T)} = x_i \mid \upsilon; T) f_{\pi_i}(\upsilon; \alpha, \beta)}{\text{pr}(X_{i(T)} = x_i)}
+\end{equation*}
+The marginal likelihood can be found in a straightforward manner
+\begin{equation*}
+\text{pr}(X_{i(T)} = x_i) = \int_0^1 \text{pr}(X_{i(T)} = x_i | \upsilon)f_{\pi_i}(\upsilon)d\upsilon = \binom{T}{x_i} \frac{\mathrm{B}(\alpha + x_i, \beta + T - x_i)}{\mathrm{B}(\alpha, \beta)}
+\end{equation*}
+Conditioning on $X_{i(T)} > 0$ results in the following truncated distribution
+\begin{equation*}
+\text{pr}(X_{i(T)} = x_i \mid X_{i(T)} > 0; \alpha, \beta, T) = \frac{\text{pr}(X_{i(T)} = x_i)}{1 - \text{pr}(X_{i(T)} = 0)} = \binom{T}{x_i} \frac{\mathrm{B}(\alpha + x_i, \beta + T - x_i)}{\mathrm{B}(\alpha, \beta) - \mathrm{B}(\alpha, \beta + T)}
+\end{equation*}
+Following the empirical Bayes approach, we estimate hyperparameters $\alpha$ and $\beta$ by maximizing the marginal likelihood.
+\begin{align*}
+    \mathcal{L}(\alpha, \beta) &= \prod_{i \in D} \binom{T}{x_i} \frac{\mathrm{B}(\alpha + x_i, \beta + T - x_i)}{\mathrm{B}(\alpha, \beta) - \mathrm{B}(\alpha, \beta + T)} \\
+    &= \prod_{i \in D} \binom{T}{x_i} \frac{\Gamma(\alpha + x_i)\Gamma(\beta + T - x_i)\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)\Gamma(\alpha + \beta + T) - \Gamma(\alpha)\Gamma(\beta + T)\Gamma(\alpha + \beta)} \\
+\end{align*}
+Using the recursive relation of the gamma function $\Gamma(x) = (x - 1)\Gamma(x - 1)$ and the fact that $T$ and $x_i$ are natural numbers, we can re-write the marginal likelihood as
+\begin{align} \label{eq:2}
+    \mathcal{L}(\alpha, \beta) &= \prod_{i \in D} \binom{T}{x_i} \frac{\Gamma(\alpha)\Gamma(\beta)\Gamma(\alpha + \beta) \prod_{j=1}^{x_i} (\alpha + x_i - j)\prod_{j=1}^{T - x_i} (\beta + T - x_i - j)}{\Gamma(\alpha)\Gamma(\beta)\Gamma(\alpha + \beta)\{\prod_{j=1}^T(\alpha + \beta + T - j) - \prod_{j=1}^T (\beta + T - j)\}} \nonumber \\
+    &= \prod_{i \in D} \binom{T}{x_i} \frac{\prod_{j=1}^{x_i} (\alpha + x_i - j)\prod_{j=1}^{T - x_i} (\beta + T - x_i - j)}{\prod_{j=1}^T(\alpha + \beta + T - j) - \prod_{j=1}^T (\beta + T - j)}
+\end{align}
+\begin{thebibliography}{1}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{Pollock et~al.(1990)Nichols, Brownie \& Hines}]{Pollock:1990}
+\textsc{Pollock, K. H., Nichols, J. D., Brownie, C. \& Hines, J. E} (1990).
+\newblock \textit{{Statistical inference for capture-recapture experiments}}.
+\newblock \textit{Wildlife Monographs} \textbf{107}, 3--97.
+
+\end{thebibliography}
 \end{document}

### 2022-03-09 _900_output/paper/paper.tex
Finish the draft of paper body

@@
 However, since we only observe $X_{i(T)} > 0$, the above distribution must be truncated at zero.
-For convenience purposes, it is also useful to partition $\mathcal{P}$ into $D = \cup_{t=1}^T S_t = \{i \in \mathcal{P}: X_{i(T)} > 0\}$ and $U = \mathcal{P} \setminus D = \{i \in \mathcal{P}: X_{i(T)} = 0\}$.
+For convenience purposes, it is also useful to partition $\mathcal{P}$ into $D = \bigcup_{t=1}^T S_t = \{i \in \mathcal{P}: X_{i(T)} > 0\}$ and $U = \mathcal{P} \setminus D = \{i \in \mathcal{P}: X_{i(T)} = 0\}$.
 Accordingly, population size can be expressed as a sum of cardinalities of $D$ and $U$, i.e.
@@
 Setting $T = 2$ and recognizing that $\sum_{i \in D} x_i = nT$ yields the following maximum likelihood estimator of $N_U$
+\begin{equation} \label{eq:0}
+\widehat{N}_U = \frac{(n - N_D)^2}{2n - N_D}
+\end{equation}
+Thus, the population size is estimated by
 \begin{equation*}
-\widehat{N}_U = \frac{(n - N_D)^2}{2n - N_D}
+\widehat{N} = N_D + \widehat{N}_U = \frac{n^2}{2n - N_D}
 \end{equation*}
-Thus, the population size is estimated by
+The above expression can be rewritten in a more familiar manner, if we introduce the set of recaptured individuals $R = S_1 \cap S_2 = \{i \in \mathcal{P}: X_{i(2)} = 2\}$.
+The size of $D = S_1 \cup S_2$ is then the sum of sample sizes at each draw minus the size of $R$
 \begin{equation} \label{eq:1}
-\widehat{N} = N_D + \widehat{N}_U = \frac{n^2}{2n - N_D}
+\widehat{N} = \frac{n^2}{2n - 2n + N_R} = \frac{n^2}{N_R}
 \end{equation}
@@
 Assume now that $\pi_i$ varies across population units meaning that \eqref{eq:1} is no longer applicable.
-Let us consider the problem from the Bayesian perspective and assume a beta prior for inclusion probabilities with hyperparameters $\alpha$ and $\beta$
+Let us consider the problem from the Bayesian perspective and assume a beta prior for inclusion probabilities with hyperparameters $\alpha > 0$ and $\beta > 0$
 \begin{equation*}
@@
 \begin{equation*}
-\text{pr}(X_{i(T)} = x_i) = \int_0^1 \text{pr}(X_{i(T)} = x_i | \upsilon)f_{\pi_i}(\upsilon)d\upsilon = \binom{T}{x_i} \frac{\mathrm{B}(\alpha + x_i, \beta + T - x_i)}{\mathrm{B}(\alpha, \beta)}
+\text{pr}(X_{i(T)} = x_i) = \int_0^1 \text{pr}(X_{i(T)} = x_i | \upsilon; T)f_{\pi_i}(\upsilon; \alpha, \beta)d\upsilon = \binom{T}{x_i} \frac{\mathrm{B}(\alpha + x_i, \beta + T - x_i)}{\mathrm{B}(\alpha, \beta)}
 \end{equation*}
@@
 \end{align*}
-Using the recursive relation of the gamma function $\Gamma(x) = (x - 1)\Gamma(x - 1)$ and the fact that $T$ and $x_i$ are natural numbers, we can re-write the marginal likelihood as
+Using the recursive relation of the gamma function $\Gamma(x) = (x - 1)\Gamma(x - 1)$ and the fact that $T$ and $x_i$ are natural numbers, we can rewrite the marginal likelihood as
 \begin{align} \label{eq:2}
@@
 \end{align}
+For inclusion probabilities $\pi_i$, it holds
+\begin{equation} \label{eq:3}
+E(\sum_{i \in \mathcal{P}} \pi_i) = N \frac{\alpha}{\alpha + \beta} = n
+\end{equation}
+The marginal likelihood \eqref{eq:2} is to be maximized subject to \eqref{eq:3}.
+Note that \eqref{eq:3} can also be viewed as centering our prior at $n / N$.
+In other words, the hyperparameters determine only the variance of the prior.
+From the constraint \eqref{eq:3}, we have $\beta = Nn^{-1}\alpha - \alpha$.
+Thus, the likelihood can be reparametrized in terms of $N_U$ and $\alpha$
+\begin{equation} \label{eq:4}
+    \mathcal{L}(\alpha, N_U) = \prod_{i \in D} \binom{T}{x_i} \frac{\prod_{j=1}^{x_i} (\alpha + x_i - j)\prod_{j=1}^{T - x_i} (Nn^{-1} \alpha - \alpha + T - x_i - j)}{\prod_{j=1}^T(Nn^{-1} \alpha + T - j) - \prod_{j=1}^T (Nn^{-1} \alpha - \alpha + T - j)}
+\end{equation}
+where $N = N_D + N_U$.
+
+The maximum likelihood estimates of $N_U$ and $\alpha$ have no closed-form expressions and must be found numerically.
+
+Setting $T = 2$ and $\alpha \to \infty$ will yield the estimator of $N_U$ equivalent to \eqref{eq:0}, as the variance of the prior tends to 0 resulting in a degenerate distribution centered at $n / N$.
+This can be shown more formally by taking the limit of \eqref{eq:4} with respect to $\alpha \to \infty$.
+
+First, partition $D$ into the set of recaptured individuals $R = S_1 \cap S_2 = \{i \in \mathcal{P}: X_{i(2)} = 2\}$ and the set of individuals captured only once $C = \{i \in \mathcal{P}: X_{i(2)} = 1\}$ with sizes $N_R$ and $N_C$ respectively.
+Then, the likelihood simplifies to
+\begin{align*}
+    \mathcal{L}(\alpha, N_U) &= \prod_{i \in R \cup C} \binom{2}{x_i} \frac{\prod_{j=1}^{x_i} (\alpha + x_i - j)\prod_{j=1}^{2 - x_i} (Nn^{-1} \alpha - \alpha + 2 - x_i - j)}{\prod_{j=1}^2(Nn^{-1} \alpha + 2 - j) - \prod_{j=1}^2 (Nn^{-1} \alpha - \alpha + 2 - j)} \\
+    &= \frac{\prod_{i \in C} 2\alpha(Nn^{-1}\alpha - \alpha + 1)\prod_{i \in R} \alpha (\alpha + 1)}{(2Nn^{-1}\alpha^2 - \alpha^2 + \alpha)^{N_C + N_R}} \\
+    &= \frac{\alpha^{N_C + N_R}(\alpha + 1)^{N_R}\{2(Nn^{-1}\alpha - \alpha + 2)\}^{N_C}}{(2Nn^{-1}\alpha^2 - \alpha^2 + \alpha)^{N_C + N_R}} \\
+    &= \left\{\frac{\alpha + 1}{\alpha(2Nn^{-1} - 1) + 1}\right\}^{N_R} \left\{\frac{2\alpha(Nn^{-1} - 1) + 4}{\alpha(2Nn^{-1} - 1) + 1}\right\}^{N_C}
+\end{align*}
+The limit of this expression as $\alpha \to \infty$ is
+\begin{align} \label{eq:5}
+&\lim_{\alpha \to \infty} \left[\left\{\frac{\alpha + 1}{\alpha(2Nn^{-1} - 1) + 1}\right\}^{N_R} \left\{\frac{2\alpha(Nn^{-1} - 1) + 4}{\alpha(2Nn^{-1} - 1) + 1}\right\}^{N_C}\right] \nonumber \\
+&= \lim_{\alpha \to \infty} \left[\left\{\frac{1 + \alpha^{-1}}{2Nn^{-1} - 1 + \alpha^{-1}}\right\}^{N_R} \left\{\frac{2(Nn^{-1} - 1) + 4\alpha^{-1}}{2Nn^{-1} - 1 + \alpha^{-1}}\right\}^{N_C}\right] \nonumber \\
+&= \frac{2^{N_C}(Nn^{-1} - 1)^{N_C}}{(2Nn^{-1} - 1)^{N_D}}
+\end{align}
+since $N_R + N_C = N_D$. By taking the logarithm of \eqref{eq:5} and maximizing with respect to $N$ we have
+\begin{equation*}
+\widehat{N} = \frac{n(2N_D - N_C)}{2(N_C - N_D)}
+\end{equation*}
+As $N_C = N_D - N_R$ and $N_D = 2n - N_R$, we arrive to
+\begin{equation*}
+\widehat{N} = \frac{n^2}{N_R}
+\end{equation*}
+which is the Lincoln-Peterson estimator for fixed $n$ as in \eqref{eq:1}. From it, the estimator \eqref{eq:0} follows.
+
+\section{Discussion}
 \begin{thebibliography}{1}

### 2022-04-25 _100_utils.jl
Add Lincoln and Schnabel estimators, trials

@@ -1,14 +1,14 @@
 module Utils
 
 import Distributions: Beta
-import SpecialFunctions: beta
+import StatsBase: addcounts!
 
 export loglh, loglh_truncated
 
 
-function monte_carlo(alpha_parameter, beta_parameter, x_i, n, mcdraws)
+function monte_carlo(alpha, beta, x_i, n, mcdraws)
     sum_term = 0.0;
-    points = rand(Beta(alpha_parameter, beta_parameter), mcdraws);
+    points = rand(Beta(alpha, beta), mcdraws);
     for p in points
         g = (n .* p).^x_i .* (1.0 .- n .* p).^(1.0 .- x_i);
         sum_term += prod(g);
@@ -16,25 +16,22 @@ function monte_carlo(alpha_parameter, beta_parameter, x_i, n, mcdraws)
     return sum_term / mcdraws;
 end
 
-function trapezoid(alpha_parameter, beta_parameter, x_i, n, draws, a, b)
-    delta = (b - a) / draws;
-    points = a:delta:b;
-    sum_term = 0.0;
-    for p in points
-        g = (n .* p).^x_i .* (1.0 .- n .* p).^(1.0 .- x_i);
-        integrand = ( prod(g) * p^(alpha_parameter - 1.0)
-                      * (1.0 - p)^(beta_parameter - 1.0)
-                      / beta(alpha_parameter, beta_parameter) );
-        if (p == a || p == b)
-            sum_term += integrand;
-        else
-            sum_term += 2.0 * integrand;
-        end
-    end
-    return delta / 2.0 * sum_term;
+function lincoln(S)
+    r = intersect(Set(S[1]), Set(S[2]));
+    return prod([length(s) for s in S]) / length(r);
 end
 
-function loglh(alpha_parameter, N_u, S, draws)
+function schnabel(S)
+    K = Dict{Int, Int}();
+    for s in S
+        addcounts!(K, s);
+    end
+    n = [length(s) for s in S];
+    T = length(S);
+    return sum(n[2:T] .* cumsum(n)[2:T]) / sum(values(K) .- 1);
+end
+
+function loglh(alpha, N_u, S, draws)
     O = Set([i for j in S for i in j]);
     N_o = length(O);
     T = length(S);
@@ -42,28 +39,25 @@ function loglh(alpha_parameter, N_u, S, draws)
     sum_term = 0.0;
     for i in O
[…]

### 2022-04-25 _110_simulate_data.jl

@@ -5,20 +5,27 @@ using SpecialFunctions
 using StatsBase
 using NLopt
 
-ALPHA = 1;
-N = 10000;
+ALPHA = 0.001;
+N = 1000;
 T = 2;
-AVG_SAMPLE_SIZE = 250;
+AVG_SAMPLE_SIZE = 25;
 
-n = rand(Poisson(AVG_SAMPLE_SIZE), T);
-d = Beta(ALPHA, ALPHA * (N - 1));
-p = rand(d, N);
-S = [sample(1:N, ProbabilityWeights(p), i, replace=false) for i in n];
-O = Set([i for j in S for i in j]);
-LL(x, grad) = -loglh_truncated(x[1], x[2], S, 100000);
-opt = Opt(:LN_SBPLX, 2);
-lower = [0.1, 2];
-opt.lower_bounds = lower;
-opt.ftol_abs = 0.01;
-opt.min_objective = LL;
-(minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
+estis = [];
+links = [];
+for trial in 1:20
+    println("***TRIAL NO $trial***")
+    n = rand(Poisson(AVG_SAMPLE_SIZE), T);
+    d = Beta(ALPHA, ALPHA * (N - 1));
+    p = rand(d, N);
+    S = [sample(1:N, ProbabilityWeights(p), i, replace=false) for i in n];
+    O = Set([i for j in S for i in j]);
+    LL(x, grad) = -loglh_truncated(x[1], x[2], S, 100000);
+    opt = Opt(:LN_SBPLX, 2);
+    lower = [0.1, 2];
+    opt.lower_bounds = lower;
+    opt.min_objective = LL;
+    opt.xtol_abs = 0.1;
+    (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
+    push!(estis, round(minx[2]) + length(O));
+    push!(links, round(lincoln(S)));
+end

### 2022-04-27 _100_utils.jl
Decrease tolerance, MC integration points

@@ -5,6 +5,15 @@ import StatsBase: addcounts!
 
 export loglh, loglh_truncated
 
+function pareto_sampling(p, n)
+    lambs = [[i, p[i] * n] for i in 1:length(p)];
+    Q = [[i[1], i[2] / sum(lambs)[2]] for i in lambs];
+    for i in 1:N
+       u = rand();
+       Q[i][2] = (u / (1 - u)) / (Q[i][2] / (1 - Q[i][2]));
+    end
+    return [Int(i[1]) for i in sort(Q, by=x -> x[2])[1:n]];
+end
 
 function monte_carlo(alpha, beta, x_i, n, mcdraws)
     sum_term = 0.0;
@@ -22,7 +31,7 @@ function lincoln(S)
 end
 
 function schnabel(S)
-    K = Dict{Int, Int}();
+    K = Dict{Float64, Int}();
     for s in S
         addcounts!(K, s);
     end

### 2022-04-27 _110_simulate_data.jl

@@ -5,27 +5,40 @@ using SpecialFunctions
 using StatsBase
 using NLopt
 
-ALPHA = 0.001;
+import StatsBase: efraimidis_a_wsample_norep!
+
+ALPHAS = [0.1];
 N = 1000;
-T = 2;
-AVG_SAMPLE_SIZE = 25;
+T = [5];
+AVG_SAMPLE_SIZE = 50.0;
+
 
 estis = [];
 links = [];
-for trial in 1:20
-    println("***TRIAL NO $trial***")
-    n = rand(Poisson(AVG_SAMPLE_SIZE), T);
-    d = Beta(ALPHA, ALPHA * (N - 1));
+schnab = [];
+for alpha in ALPHAS
+    d = Beta(alpha, alpha * (N - 1));
     p = rand(d, N);
-    S = [sample(1:N, ProbabilityWeights(p), i, replace=false) for i in n];
-    O = Set([i for j in S for i in j]);
-    LL(x, grad) = -loglh_truncated(x[1], x[2], S, 100000);
-    opt = Opt(:LN_SBPLX, 2);
-    lower = [0.1, 2];
-    opt.lower_bounds = lower;
-    opt.min_objective = LL;
-    opt.xtol_abs = 0.1;
-    (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
-    push!(estis, round(minx[2]) + length(O));
-    push!(links, round(lincoln(S)));
+    for t in T
+        n = rand(Poisson(AVG_SAMPLE_SIZE), t);
+        for trial in 1:100
+            println("***TRIAL NO $trial, $alpha, $t***")
+            S = [efraimidis_a_wsample_norep!(1:N, Weights(p), zeros(i)) for i in n];
+            O = Set([i for j in S for i in j]);
+            LL(x, grad) = -loglh_truncated(x[1], x[2], S, 1000);
+            opt = Opt(:LN_SBPLX, 2);
+            lower = [0.01, 0];
+            opt.lower_bounds = lower;
+            opt.min_objective = LL;
+            opt.xtol_abs = 0.5;
+            (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
+            push!(estis, [round(minx[2]) + length(O), alpha, t]);
+            if t == 2
+                push!(links, [round(lincoln(S)), alpha, t]);
+            end
+            if t > 2
+                push!(schnab, [round(schnabel(S)), alpha, t]);
+            end
+        end
+    end
 end

### 2022-04-27 _100_utils.jl
Added skeleton for Chao estimator

@@ -10,7 +10,7 @@ function pareto_sampling(p, n)
     Q = [[i[1], i[2] / sum(lambs)[2]] for i in lambs];
     for i in 1:N
        u = rand();
-       Q[i][2] = (u / (1 - u)) / (Q[i][2] / (1 - Q[i][2]));
+       Q[i][2] = (u / (1.0 - u)) / (Q[i][2] / (1.0 - Q[i][2]));
     end
     return [Int(i[1]) for i in sort(Q, by=x -> x[2])[1:n]];
 end
@@ -37,7 +37,29 @@ function schnabel(S)
     end
     n = [length(s) for s in S];
     T = length(S);
-    return sum(n[2:T] .* cumsum(n)[2:T]) / sum(values(K) .- 1);
+    return sum(n[2:T] .* cumsum(n)[2:T]) / sum(values(K) .- 1.0);
+end
+
+function chao(S)
+    K = Dict{Float64, Int}();
+    for s in S
+        addcounts!(K, s);
+    end
+    f = countmap(values(K));
+    O = Set([i for j in S for i in j]);
+    return length(O) + f[1]^2.0 / (2.0 * f[2]);
+end
+
+function chao_corrected(S)
+    K = Dict{Float64, Int}();
+    for s in S
+        addcounts!(K, s);
+    end
+    f = countmap(values(K));
+    O = Set([i for j in S for i in j]);
+    m1 = 2.0 * f[2] / f[1];
+    m2 = 6.0 * f[3] / f[1];
+    return length(O) + f[1]^2.0 / (2.0 * f[2]) * (1.0 - m1 / length(S)) / (1.0 - m2 / (length(S) * m1));
 end
 
 function loglh(alpha, N_u, S, draws)

### 2022-04-27 _100_utils.jl
Adds jackknife estimator, refactor simulation code

@@ -25,48 +25,59 @@ function monte_carlo(alpha, beta, x_i, n, mcdraws)
     return sum_term / mcdraws;
 end
 
-function lincoln(S)
+function lincoln(S, n)
     r = intersect(Set(S[1]), Set(S[2]));
-    return prod([length(s) for s in S]) / length(r);
+    return prod(n) / length(r);
 end
 
-function schnabel(S)
-    K = Dict{Float64, Int}();
-    for s in S
-        addcounts!(K, s);
-    end
-    n = [length(s) for s in S];
-    T = length(S);
+function schnabel(n, T, K)
     return sum(n[2:T] .* cumsum(n)[2:T]) / sum(values(K) .- 1.0);
 end
 
-function chao(S)
-    K = Dict{Float64, Int}();
-    for s in S
-        addcounts!(K, s);
-    end
-    f = countmap(values(K));
-    O = Set([i for j in S for i in j]);
-    return length(O) + f[1]^2.0 / (2.0 * f[2]);
+function chao(N_o, f)
+    return N_o + get(f, 1, 0)^2.0 / (2.0 * get(f, 2, 0));
 end
 
-function chao_corrected(S)
-    K = Dict{Float64, Int}();
-    for s in S
-        addcounts!(K, s);
-    end
-    f = countmap(values(K));
-    O = Set([i for j in S for i in j]);
-    m1 = 2.0 * f[2] / f[1];
-    m2 = 6.0 * f[3] / f[1];
-    return length(O) + f[1]^2.0 / (2.0 * f[2]) * (1.0 - m1 / length(S)) / (1.0 - m2 / (length(S) * m1));
+function chao_corrected(N_o, T, f)
+    m1 = 2.0 * get(f, 2, 0)  / get(f, 1, 0);
+    m2 = 6.0 * get(f, 3, 0) / get(f, 1, 0);
+    return (N_o
+            + get(f, 1, 0)^2.0 / (2.0 * get(f, 2, 0))
+            * (1.0 - m1 / T) / (1.0 - m2 / (T * m1)));
 end
 
-function loglh(alpha, N_u, S, draws)
-    O = Set([i for j in S for i in j]);
+function jackknife(N_o, T, f, k)
+    if k == 1
+        return N_o + (T - 1.0) / T * get(f, 1, 0);
+    elseif k == 2
+        return (N_o
[…]

### 2022-04-27 _110_simulate_data.jl

@@ -7,37 +7,49 @@ using NLopt
 
 import StatsBase: efraimidis_a_wsample_norep!
 
-ALPHAS = [0.1];
+ALPHAS = [0.1, 0.5, 1.0, 5.0, 20.0];
 N = 1000;
-T = [5];
+T = [2, 5, 8, 10];
 AVG_SAMPLE_SIZE = 50.0;
 
 
+    
 estis = [];
 links = [];
 schnab = [];
+chaos = [];
+chaos_corr = [];
+jks = [];
 for alpha in ALPHAS
     d = Beta(alpha, alpha * (N - 1));
     p = rand(d, N);
     for t in T
         n = rand(Poisson(AVG_SAMPLE_SIZE), t);
-        for trial in 1:100
+        for trial in 1:1
             println("***TRIAL NO $trial, $alpha, $t***")
             S = [efraimidis_a_wsample_norep!(1:N, Weights(p), zeros(i)) for i in n];
             O = Set([i for j in S for i in j]);
-            LL(x, grad) = -loglh_truncated(x[1], x[2], S, 1000);
+            K = Dict{Float64, Int}();
+            for s in S
+                addcounts!(K, s);
+            end
+            f = countmap(values(K));
+            LL(x, grad) = -loglh_truncated(x[1], x[2], S, O, t, n, 1000);
             opt = Opt(:LN_SBPLX, 2);
             lower = [0.01, 0];
             opt.lower_bounds = lower;
             opt.min_objective = LL;
-            opt.xtol_abs = 0.5;
+            opt.xtol_abs = 0.1;
             (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
             push!(estis, [round(minx[2]) + length(O), alpha, t]);
+            push!(chaos, [round(chao(length(O), f)), alpha, t]);
+            push!(chaos_corr, [round(chao_corrected(length(O), t, f)), alpha, t]);
+            push!(jks, vcat([round(jackknife(length(O), t, f, k)) for k in 1:5], [alpha, t]));
             if t == 2
-                push!(links, [round(lincoln(S)), alpha, t]);
+                push!(links, [round(lincoln(S, n)), alpha, t]);
             end
             if t > 2
-                push!(schnab, [round(schnabel(S)), alpha, t]);
+                push!(schnab, [round(schnabel(n, t, K)), alpha, t]);
             end
         end
     end

### 2022-04-28 _110_simulate_data.jl
Lots of stuff

@@ -6,26 +6,39 @@ using StatsBase
 using NLopt
 
 import StatsBase: efraimidis_a_wsample_norep!
+import Random: seed!
 
 ALPHAS = [0.1, 0.5, 1.0, 5.0, 20.0];
 N = 1000;
-T = [2, 5, 8, 10];
-AVG_SAMPLE_SIZE = 50.0;
+T = [2, 4, 10];
+AVG_SAMPLE_SIZE = 75.0;
 
 
-    
+function write_row(filename, row)
+    open(filename, "a") do io
+        for i in row
+            if i != row[length(row)]
+                print(io, i, ",");
+            else
+                print(io, i, "\n");
+            end
+        end
+    end
+end
+
 estis = [];
 links = [];
 schnab = [];
 chaos = [];
 chaos_corr = [];
 jks = [];
+seed!(777);
 for alpha in ALPHAS
     d = Beta(alpha, alpha * (N - 1));
     p = rand(d, N);
     for t in T
         n = rand(Poisson(AVG_SAMPLE_SIZE), t);
-        for trial in 1:1
+        for trial in 1:75
             println("***TRIAL NO $trial, $alpha, $t***")
             S = [efraimidis_a_wsample_norep!(1:N, Weights(p), zeros(i)) for i in n];
             O = Set([i for j in S for i in j]);
@@ -41,15 +54,21 @@ for alpha in ALPHAS
             opt.min_objective = LL;
             opt.xtol_abs = 0.1;
             (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
-            push!(estis, [round(minx[2]) + length(O), alpha, t]);
+            push!(estis, [minx[1] ,round(minx[2]) + length(O), alpha, t]);
+            write_row("estis.csv", [minx[1] ,round(minx[2]) + length(O), alpha, t]);
             push!(chaos, [round(chao(length(O), f)), alpha, t]);
+            write_row("chaos.csv", [round(chao(length(O), f)), alpha, t]);
             push!(chaos_corr, [round(chao_corrected(length(O), t, f)), alpha, t]);
+            write_row("chaos_corr.csv", [round(chao_corrected(length(O), t, f)), alpha, t]);
             push!(jks, vcat([round(jackknife(length(O), t, f, k)) for k in 1:5], [alpha, t]));
+            write_row("jks.csv", vcat([round(jackknife(length(O), t, f, k)) for k in 1:5], [alpha, t]));
             if t == 2
[…]

### 2022-04-28 _120_simulate_graph.jl
Lots of stuff

@@ -1,0 +1,51 @@
+using .Utils
+
+using Graphs
+using StatsBase
+using NLopt
+
+import Random: seed!
+
+NODES = 1000;
+EDGES_PER_NODE = 4;
+T = [2, 4, 10];
+SNOWBALL_WAVES = 2;
+
+function sample_triangles(g, waves, root)
+    gs = egonet(g, root, waves);
+    s_tri = Set([Set(i) for i in simplecycles_limited_length(gs, 3) if length(i) == 3]);
+    return s_tri;
+end
+
+function write_row(filename, row)
+    open(filename, "a") do io
+        for i in row
+            if i != row[length(row)]
+                print(io, i, ",");
+            else
+                print(io, i, "\n");
+            end
+        end
+    end
+end
+
+
+G = barabasi_albert(NODES, EDGES_PER_NODE);
+estis = [];
+for t in T
+    for trial in 1:50
+        roots = rand(1:NODES, t);
+        S = [sample_triangles(G, SNOWBALL_WAVES, r) for r in roots];
+        n = [length(s) for s in S];
+        O = Set([i for j in S for i in j]);
+        LL(x, grad) = -loglh_truncated(x[1], x[2], S, O, t, n, 1000);
+        opt = Opt(:LN_SBPLX, 2);
+        lower = [0.01, 0];
+        opt.lower_bounds = lower;
+        opt.min_objective = LL;
+        opt.xtol_abs = 0.1;
+        (minf, minx, ret) = NLopt.optimize(opt, [5.0, 10.0]);
+        push!(estis, [minx[1] ,round(minx[2]) + length(O), t]);
+        write_row("tris.csv", [minx[1] ,round(minx[2]) + length(O), t]);
+    end
+end

### 2022-07-25 simulation_results.Rmd
Add new plots and summary of simulations

---
title: "Comparison with other estimators in simulation"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A simulation study is conducted to assess the performance of our estimator in 
comparison to the existing methods in the literature. I differentiate between 
four cases: equal inclusion probabilities with equal sample sizes, equal 
inclusion probabilities with different sample sizes, unequal probabilities with 
equal sample sizes and unequal probabilities with different sample size.

For each case, the simulation is conducted as follows. First, the true population
size $N$ is fixed along with the $\alpha$ parameter. Then, $p_i$ are generated 
for each individual $i = 1, \dots, N$ by drawing $N$ values from $Beta(\alpha, \alpha(N - 1))$. 
For cases with different sample sizes, a set of $T$ sample sizes $\{n_1, \dots, n_T\}$ 
is generated randomly, while for cases with equal sample sizes $n_1 = n_2 = ... = n_T = n$ 
for some fixed $n$. Finally, $T$ samples are generated by drawing $n_t$ times from 
population without replacement in accordance with inclusion probabilities 
$\pi_{it} = n_t \cdot pi_i$ for each $t = 1, \dots, T$ and $i = 1, \dots, N$.

Given the simulated samples, the estimates of the population size are computed. 
The above procedure is repeated 100 times for each case and the resulting population 
size estimates are illustrated in a boxplot for comparison with other approaches.

Since the loglikelihood function involves integrals without analytical solutions, 
the optimization is performed numerically. The integrals are approximated by 
Monte Carlo integration, as we can easily produce draws from the beta distribution. 
Another advantage of solving the integrals by Monte Carlo integration, as opposed 
to other deterministic numerical integration methods, is that there is no need 
to evaluate beta functions, thus avoiding numerical overflow. For each loglikelihood 
evaluation, the Monte Carlo integration is performed with 1000 draws for each 
integral term.

The loglikelihood is fitted using Rowan (1990)'s Subplex optimization algorithm, which 
is a more robust and efficient variant of Nelder-Mead Simplex method.

\section{Equal probabilities}
For equal inclusion probabilities, $\alpha \to \infty$ and $T$ samples are drawn 
via simple random sampling without replacement. For the next two cases, the 
population size is set to $N = 1000$ with $T \in \{2, 4, 10\}$. For $T = 2$, only 
simulations with recaptures are considered.

\subsection{Equal sample sizes}
For this case, the sample sizes were fixed at $n_1 = ... = n_T = 30$.

The boxplots below describe the distribution of estimates for various values of 
$T$ for different estimators. To control for variability of estimates, a reciprocal 
was taken. The blue vertical line represents true value at $1 / N = 0.001$.

![](./_900_output/figures/eqp_eqn/estimates_boxplots.pdf)

Additionally, similar boxplots were produced separately comparing our estimator 
with the jackknife capture-recapture estimator for different values of the 
jackknife's hyperparameter $k$.

![](./_900_output/figures/eqp_eqn/jks_boxplots.pdf)

The boxplots of estimates of the $\alpha$ parameter are also given below. As the 
true $\alpha$ is infinity in this case, the estimated values are accordingly large 
and so the x-axis is given in log scale.

![](./_900_output/figures/eqp_eqn/alpha_hat.pdf)

\subsection{Unequal sample sizes}
For the case of equal inclusion probabilities with different sample sizes, the 
simulation was conducted with the same parameters as previously but with sample 
sizes generated from $n_t \sim Poisson(30)$ for all $t = 1, \dots, T$, producing 
the same set of figures.

![](./_900_output/figures/eqp_eqn/estimates_boxplots.pdf)

![](./_900_output/figures/eqp_eqn/jks_boxplots.pdf)

![](./_900_output/figures/eqp_eqn/alpha_hat.pdf)

\section{Unequal probabilities}
The simulation of unequal probability sampling is performed with the Sampford 
algorithm. Recall that parameter $\alpha$ determines the heterogeneity of inclusion 
probabilities across population. For $\alpha \to \infty$, all population units 
have the same inclusion probability. Conversely, smaller values of $\alpha$ 
correspond to populations with highly heterogeneous inclusion probabilities. 
For this reason, the simulation is performed for different values of $\alpha$, 
namely $\alpha \in \{0.1, 3, 10\}$.

To get a better sense of accuracy of estimators with increasing data, the 
sample draws are changed to $T \in \{2, 5, 10, 15, 20\}$.

\subsection{Equal sample sizes}
Unlike the case of equal probability sampling where $\pi_{it} = 1 / N$, in 
unequal probability sampling,  the assumption $\pi_{it} = n_t \cdot p_i$ must be 
ensured, which translates to enforcing $0 \leq \pi_{it} \leq 1$. For fixed 
$n_1 = ... = n_T = n$ and $\alpha \in \{0.1, 3, 10\}$, the population size $N$ 
must be chosen appropriately.

Using model assumptions,

$$
\text{pr}(\pi_{it} \leq 1) = \text{pr}(np_i \leq 1) = \text{pr}(p_i \leq 1 / n)
= I_{1/n}(\alpha, \alpha(N - 1))
$$
where $I_{x}(\alpha, \beta)$ is the regularized incomplete beta function.

Following assumption of independence of inclusion probabilities,

$$
\text{pr}(p_i \leq 1 / n \ \forall i = 1, \dots, N) = \prod_{i=1}^N I_{1/n}(\alpha, \alpha(N - 1)) 
= [I_{1/n}(\alpha, \alpha(N - 1))]^N
$$
Thus, for fixed $\alpha$ and $n$, the population size $N$ must be chosen such 
that $\text{pr}(p_i \leq 1 / n \ \forall i = 1, \dots, N)$ is high enough. For 
this case, I set $n = 30, N = 3000$, therefore

$$
\text{pr}(p_i \leq 1 / 30 \ \forall i = 1, \dots, 3000) \approx 1.0 \quad \forall \alpha \in \{0.1, 3, 20\} 
$$

\subsubsection{$\alpha = 0.1$}

![](./_900_output/figures/diffp_eqn/estimates_boxplots_0.1.pdf)

![](./_900_output/figures/diffp_eqn/jks_boxplots_0.1.pdf)

![](./_900_output/figures/diffp_eqn/alpha_hat_0.1.pdf)

Finally, for reference, the plots below show the loglikelihood function for 
different values of each parameter while fixing the other at the true value for 
each $T$ and simulation replication.

![](./_900_output/figures/diffp_eqn/alpha_trace_2_0.1.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_5_0.1.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_10_0.1.pdf)
![](./_900_output/figures/diffp_eqn/alpha_trace_15_0.1.pdf)
![](./_900_output/figures/diffp_eqn/alpha_trace_20_0.1.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_2_0.1.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_5_0.1.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_10_0.1.pdf)
![](./_900_output/figures/diffp_eqn/Nu_trace_15_0.1.pdf)
![](./_900_output/figures/diffp_eqn/Nu_trace_20_0.1.pdf)

\newpage

\subsubsection{$\alpha = 3$}

![](./_900_output/figures/diffp_eqn/estimates_boxplots_3.0.pdf)

![](./_900_output/figures/diffp_eqn/jks_boxplots_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_hat_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_2_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_5_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_10_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_15_3.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_20_3.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_2_3.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_5_3.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_10_3.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_15_3.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_20_3.0.pdf)

\newpage

\subsubsection{$\alpha = 10$}

![](./_900_output/figures/diffp_eqn/estimates_boxplots_10.0.pdf)

![](./_900_output/figures/diffp_eqn/jks_boxplots_10.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_hat_10.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_2_0.1.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_5_10.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_10_10.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_15_10.0.pdf)

![](./_900_output/figures/diffp_eqn/alpha_trace_20_10.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_2_10.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_5_10.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_10_10.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_15_10.0.pdf)

![](./_900_output/figures/diffp_eqn/Nu_trace_20_10.0.pdf)

## References
T. Rowan, "Functional Stability Analysis of Numerical Algorithms", Ph.D. thesis, 
Department of Computer Sciences, University of Texas at Austin, 1990.

### 2022-07-25 _105_plot_likelihood.jl

@@ -6,12 +6,10 @@ using Distributions
 
 import Random: seed!
 
-T = [2, 5, 10, 20];
-AVERAGE_SAMPLE_SIZE = 30;
+T = [2, 5, 10, 15, 20];
+AVG_SAMPLE_SIZE = 30;
 ALPHAS = [0.1, 3.0, 10.0];
 N = 3000;
-ALPHA_PLOT_RANGE = 0.1:1:25.1;
-Nu_PLOT_RANGE = 0:100:5100;
 
 seed!(111);
 for alpha in ALPHAS
@@ -25,9 +23,25 @@ for alpha in ALPHAS
     end
     for t in T
         S = samples[1:t]
-        println("***TRIAL NO $alpha, $trial, $t***")
         O = Set([i for j in S for i in j]);
-        L(a, Nu) = loglh_truncated(a, Nu, S, O, t, n[1:t], 1000);
-        plot(ALPHA_PLOT_RANGE, Nu_PLOT_RANGE, L, st =:contourf, layout = l);
-        xlabel!("Alpha")
-        ylabel!("N_u")
+        L(a, Nu) = loglh_truncated(a, Nu, S, O, t, n[1:t], 5000);
+        plot(0.1:1:(alpha + 2.1),
+             0:100:((N - length(O)) + 500),
+             L,
+             st =:contourf,
+             title="Loglikelihood surface for T = $t and alpha = $alpha",
+             size=(1024, 640));
+        xlabel!("Alpha");
+        ylabel!("N_u");
+        savefig("contour_$(t)_$(alpha).pdf")
+        plot(0.1:1:(alpha + 2.1),
+             0:100:((N - length(O)) + 500),
+             L,
+             st =:surface,
+             title="Loglikelihood surface for T = $t and alpha = $alpha",
+             size=(1024, 640));
+        xlabel!("Alpha");
+        ylabel!("N_u");
+        savefig("surface_$(t)_$(alpha).pdf")
+    end
+end

### 2022-07-25 _130_diffprobs_eqsamples.jl

@@ -12,7 +12,7 @@ import Random: seed!
 N = 3000;
 T = [2, 5, 10, 15, 20];
 TRIALS = 100;
-ALPHAS = [0.1, 3, 10];
+ALPHAS = [0.1];
 AVG_SAMPLE_SIZE = 30;
 
 
@@ -42,12 +42,13 @@ for alpha in ALPHAS
                 addcounts!(K, s);
             end
             f = countmap(values(K));
-            LL(x, grad) = -loglh_truncated(x[1], x[2], S, O, t, n[1:t], 1000);
+            LL(x, grad) = -mean([loglh_truncated(x[1], x[2], S, O, t, n[1:t], 100) for i in 1:100]);
             opt = Opt(:LN_SBPLX, 2);
             lower = [0.01, 0];
             opt.lower_bounds = lower;
             opt.min_objective = LL;
-            opt.xtol_abs = 0.1;
+            opt.xtol_abs = 0.5;
+            opt.maxeval = 100;
             (minf, minx, ret) = NLopt.optimize(opt, [5.0, length(O)]);
 #             LL(x) = -loglh_truncated(x[1], x[2], S, O, t, n[1:t], 3000);
 #             x0 = [1.0, length(O)];

### 2022-07-25 plot_estimates.py

@@ -114,11 +114,11 @@ for alpha in ALPHAS:
         fig, ax = plt.subplots(figsize=(18,12))
         cut = nu_trace[nu_trace["T"] == t]
         for row in range(cut.shape[0]):
-            ax.plot(np.log(NU_RANGE[1:]), cut.iloc[row, 1:len(NU_RANGE)]) # Discard first values to take logs
+            ax.plot(NU_RANGE[1:], cut.iloc[row, 1:len(NU_RANGE)]) # Discard first values to take logs
             ax.axvline(x=cut["true_Nu"].iloc[t], color="gold", alpha=0.1)
             ax.set_ylabel("Loglikelihood")
             ax.set_xlabel(r"$N_U$")
-            ax.set_title(r"Loglikelihood for different values of $N_U$ at $\alpha = \hat{\alpha}_{ML}$ for each trial. Vertical line shows true $N_U$")
+            ax.set_title(r"Loglikelihood for different values of $N_U$ at $\hat{\alpha} = \alpha$ for each trial. Vertical line shows true $N_U$," + " T = {}.".format(t))
         fig.savefig(FIGS.format("Nu_trace_{}_{}.pdf".format(t, alpha)),
                     bbox_inches="tight")
         plt.close(fig)
@@ -134,11 +134,12 @@ for alpha in ALPHAS:
     for t in T:
         fig, ax = plt.subplots(figsize=(18,12))
         cut = alpha_trace[alpha_trace["T"] == t]
+        ax.axvline(x=alpha, color="gold")
         for row in range(cut.shape[0]):
-            ax.plot(np.log(ALPHA_RANGE), cut.iloc[row, 1:len(ALPHA_RANGE) + 1])
+            ax.plot(ALPHA_RANGE, cut.iloc[row, 1:len(ALPHA_RANGE) + 1])
             ax.set_ylabel("Loglikelihood")
             ax.set_xlabel(r"$\alpha$")
-            ax.set_title(r"Loglikelihood for different values of $\alpha$ at $N_U = \hat{N}_{U(ML)}$ for each trial.")
+            ax.set_title(r"Loglikelihood for different values of $\alpha$ at $\hat{N}_U = N_{U}$ for each trial. Vertical line shows true $\alpha$," + " T = {}.".format(t))
         fig.savefig(FIGS.format("alpha_trace_{}_{}.pdf".format(t, alpha)),
                     bbox_inches="tight")
         plt.close(fig)  
